Type Flattening Obfuscation

### 2020-01-31 paper/tool.tex
update

	% We show that the techniques used for type reconstruction can be bypassed,
	% and we claim that the bypass is always possible given \emph{semantics gap} between
	% type systems of high-level and machine language.
	% \keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{Introduction}
An important step in binary code \emph{decompilation}~\cite{cifuentes_reverse_1994} is to
detect high-level objects (e.g. functions, variables) from low-level machine dependent
objects~\cite{balakrishnan_divine_2007} (e.g. registers, raw memory accesses, machine
instructions) before annotate them with types of the "assumed" original language.
In a statically typed language like C, the compiler does not preserve the high-level
type information in the generated machine code,
% Because most of information about high-level types is lost in compilation,
% the machine code
then the \emph{type reconstruction} (or type recovery) from machine code requires
special techniques, a survey up until 2015 can be referenced in~\cite{caballero_type_2016}.

@@

\emph{First}, type obfuscation is unsurprisingly a side effect of data or control flow obfuscation.
Indeed, type reconstruction algorithms need data and control flow to build type constraints,
if any of them is hidden
then the algorithms cannot work correctly. Or if the function boundary is not found
(because of anti-disassembing tricks, for example), then high-level objects cannot be recognized.
\emph{Second}, high-level types seem too coarse to be worthy
of being protected, in many cases just knowing certain values of the input which make the program
exploitable is enough. But knowledge about types expands attack surfaces because more analysis
can be proceeded, beside decompilation see the survey~\cite{caballero_type_2016} for a
more complete list.
% We exploit the \emph{semantics gap}

\section{Type flattening compiler}
We present \emph{uCc}, an open source C compiler that explicitly obfuscates high-level types, \emph{uCc}'s
objective is to make reconstructing C types from machine codes harder.
The compiler is implemented in Rust, the source code and a brief user guide are given at~\cite{ta_ucc_nodate};
different from other code obfuscation implementations, \emph{uCc} uses cranelift~\cite{noauthor_cranelift_nodate}
as backend.

@@

We test outputs of \emph{uCc} against some decompilers, table~\ref{description} describes briefly
tested functions and their types in source codes, table~\ref{flattening} shows results of
decompilers on binaries generated by \emph{uCc}. The effect of obfuscation is to make the types
of functions harder to be reconstructed. Indeed, the compilers know very little about the original
types. Basically, any argument in the function signature is simply assigned by either \texttt{int64}
or \texttt{uint64}, the highest types in the \emph{conversion rank} of integer types, we call this
effect \emph{type flattening}.

It may worth to note that the \emph{type flattening} is given by \emph{uCc}, not by the source code
complexity of tested functions. Indeed if they are compiled by, \emph{clang} or \emph{gcc} for example,
then their signatures are reconstructed precisely
by the decompilers. Last but not least, \emph{uCc} attacks directly on type reconstruction algorithms,
it does not use obfuscation tricks
% , for example making disassembling or function boundary detection hard,
which make side effects on type
reconstruction: functions in tested binaries are perfectly detectable, disassemblable and
decompilable.

@@

>>>>

	% We show that the techniques used for type reconstruction can be bypassed,
	% and we claim that the bypass is always possible given \emph{semantics gap} between
	% type systems of high-level and machine language.
	% \keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{Introduction}
An important step in binary code \emph{decompilation}~\cite{cifuentes_reverse_1994} is to
detect high-level objects (e.g. functions, variables) from low-level machine dependent
objects~\cite{balakrishnan_divine_2007} (e.g. registers, raw memory accesses, machine
instructions) before annotate them with types of the "assumed" original language.
In a statically typed language like C, the compiler does not preserve the high-level
type information in the generated machine code,
% Because most of information about high-level types is lost in compilation,
% the machine code
then the \emph{type reconstruction} (or type recovery) from machine code requires
special techniques, a survey up until 2015 can be referenced in~\cite{caballero_type_2016}, more
recent research includes~\cite{noonan_polymorphic_2016, maier_typeminer_2019, katz_estimating_2016}.

@@

\emph{First}, type obfuscation is unsurprisingly a side effect of data or control flow obfuscation.
Indeed, type reconstruction algorithms need data and control flow to build type constraints,
if any of them is hidden
then the algorithms cannot work correctly. Or if the function boundary is not found
(because of anti-disassembly tricks, for example), then high-level objects cannot be recognized.
\emph{Second}, high-level types seem too coarse to be worthy
of being protected, in many cases just knowing certain values of the input which make the program
exploitable is enough. But knowledge about types expands attack surfaces because more analysis
can be proceeded, beside decompilation see the survey~\cite{caballero_type_2016} for a
more complete list.
% We exploit the \emph{semantics gap}

\subsection*{Type flattening obfuscation}
Under our interest, the studies in binary code type reconstruction can be roughly divided into
two directions: one is based on semantics of programs, the other is not. We focus on the former,
and also on methods using static analysis. The research in this vein starts with
Mycroft~\cite{mycroft_type-based_1999}, and improved in~\cite{lee_tie_2011,noonan_polymorphic_2016,elwazeer_scalable_2013}

\section{Compiler for type flattening}
We present \emph{uCc}, an open source C compiler that explicitly obfuscates high-level types, \emph{uCc}'s
objective is to make reconstructing C types from machine codes harder.
The compiler is implemented in Rust, the source code and a brief user guide are given at~\cite{ta_ucc_nodate};
different from other code obfuscation implementations, \emph{uCc} uses cranelift~\cite{noauthor_cranelift_nodate}
as backend.

@@

\subsection*{Evaluation}
We test outputs of \emph{uCc} against some decompilers, table~\ref{description} describes briefly
tested functions and their types in source codes, table~\ref{flattening} shows results of
decompilers on binaries generated by \emph{uCc}. The effect of obfuscation is to make the types
of functions harder to be reconstructed. Indeed, the decompilers know very little about the original
types. Basically, any argument in the function signature is simply assigned by either \texttt{int64}
or \texttt{uint64}, the highest types in the \emph{conversion rank} of integer types, we call this
effect \emph{type flattening}.

@@

It may worth to note that the \emph{type flattening} is given by \emph{uCc}, not by the source code
complexity of tested functions. Indeed if they are compiled by, \emph{clang} or \emph{gcc} for example,
then their signatures are reconstructed precisely
by the decompilers. Last but not least, \emph{uCc} attacks directly on type reconstruction algorithms,
it doesn't use obfuscation techniques
% , for example making disassembling or function boundary detection hard,
whose side effects affect the type
reconstruction. Indeed, functions in tested binaries are perfectly detectable, disassemblable and
decompilable; only their decompiled types are in question.
